---
title: "Project 1"
author: "Richard Wyllia"
date: "1/3/2023"
output: html_document
---
 
```{r setup, include=FALSE}
knitr:: opts_chunk$set(echo = TRUE, warning = FALSE, fig.width = 10, fig.height = 5, fig.keep = "all")
path = getwd()
library(ggplot2)
```

##### Gathering Data on Steps
```{r Processing data}
activity<- read.csv("activity.csv")  
  activity$date<- as.POSIXct(activity$date, "%Y%m%d")
  day<- weekdays(activity$date)

activity<- cbind(activity, day)

summary(activity)
```

```{r Chart}
activityTotalSteps<- with(activity, aggregate(steps, by = list(date), sum, na.rm = TRUE))

names(activityTotalSteps)<- c("Date","Steps")

totalStepsdf<- data.frame(activityTotalSteps)
g<- ggplot(totalStepsdf, aes(x = Steps)) + geom_histogram(breaks = seq(0, 25000, by = 2500), fill = "green", col = "black") + ylim(0, 30) + xlab("Total Steps Taken Per Day") + ylab("FREQUENCY") + ggtitle("Total Number of Steps Taken on a Day") + theme_classic(base_family = "serif")
print(g)
```

```{r Mean and Median}
mean(activityTotalSteps$Steps)
median(activityTotalSteps$Steps)
```
### Q1.) The Mean of the total number of steps taken per day is 9354.23 and the median is 10,395 per day.

```{r Daily Average}
averageDailyActivity<- aggregate(activity$steps, by = list(activity$interval), FUN = mean, na.rm = TRUE)

names(averageDailyActivity)<- c("Interval", "Mean")
averageActivitydf<- data.frame(averageDailyActivity)
j<- ggplot(averageActivitydf, mapping = aes(Interval, Mean)) + geom_line(col = "green") + xlab("Interval") + ylab("Average Number of Steps") + ggtitle("Average Steps per Interval") + theme_classic(base_family = "serif")
averageDailyActivity[which.max(averageDailyActivity$Mean), ]$Interval 
print(j)
```
### Q2.) The 835 mark has the most steps on average out of all 5 minute intervals
```{r Missing Values}
sum(is.na(activity$steps))

imputedSteps<- averageDailyActivity$Mean[match(activity$interval, averageDailyActivity$Interval)]

activityImputed<- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputedSteps, no = activity$steps))

totalActivityImputed<- aggregate(steps ~ date, activityImputed, sum)

names(totalActivityImputed)<- c("date", "dailySteps")
```

```{r}
totalImputedStepsdf<- data.frame(totalActivityImputed)

p<- ggplot(totalImputedStepsdf, aes(x = dailySteps)) +
  geom_histogram(breaks = seq(0, 25000, by = 2500), fill = "blue",
  col = "black") + ylim(0, 30) +
  xlab("Total Steps Taken Per Day") +
  ylab("Frequency") +
  ggtitle("Total Number of Steps Taken Daily") +
  theme_classic(base_family = "serif")

print(p)

mean(totalActivityImputed$dailySteps)
median(totalActivityImputed$dailySteps)
```
### Q3.)The mean and median were the same at 10,766.19 steps taken per day.

```{r}
activity$date<- as.Date(strptime(activity$date, format = "%Y-%m-%d"))

activity$dayType<- sapply(activity$date,function(x)
  {if(weekdays(x)== "Saturday"| weekdays(x)== "Sunday")
  {y<- "Weekend"}
  else{y<- "Weekday"}
  y})

activityByDay<- aggregate(steps ~ interval + dayType, activity, mean, na.rm = TRUE)
```


dayPlot<- ggplot(activityByDay, aes(x = interval, y = Steps, color = dayType)) +
  geom_line() + ggtitle("Average Daily Steps by Day Type") +
  xlab("Interval") +
  ylab("Average Number of Steps") +
  facet_wrap(~dayType, ncol = 1, nrow = 2) + 
  scale_color_discrete(name = "Day Type") +
  theme_classic(base_family = "serif")
print(dayPlot)


